\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2013-11-29 delays.tex
Updated the paper

Figure~\ref{fig:5} pictorially represents the semantics of these delay
constructs. The construct $delay (M..N)$ requires that once executed the
control-flow suspends execution for \textit{at\_least} $M$ units and
\textit{at\_most} $N$ units, depicted as bounded box in
Figure~\ref{fig:5}. There is no upper bound for the second delay form
$delay(M)$, which requires that the control-flow suspends for
\textit{at\_least} $M$ units, but the following statement can be
executed after some countable delay, depicted with the unbounded box in
Figure~\ref{fig:5}. Finally, $delay(M..M)$, requires that there is an
\textit{exact} delay of $M$ units, i.e., the control-flow suspends
execution for \textit{at\_least} and \textit{at\_most} $M$ units.

@@

This algorithm results in the \underline{smallest} relaxation required
for the real-time delay to be satisfied. The algorithm takes as input
the WCRT and set $S_2$, recall that set $S_2$ represents the logical
ticks required to satisfy the real-time requirement at the BCRT. We take
the very first value from set $S_2$ and multiply it with the WCRT to get
the relaxation $N$. The first element of set $S_2$ is returned as the
logical tick delay \texttt{d}. The fundamental observation is that we
have to delay for a minimum of $M$ units of real-time, hence,
under-approximation is out of question. We can still over-approximate,
but to reduce the resultant error, we should over-approximate by least
possible value, which is the lower bound of set $S_2$. Thus, the lower
bound is considered to be the only element shared between the two sets
$S_1$ and $S_2$ and accordingly, the upper real-time bound is relaxed by
the multiplication of WCRT and the first element of set $S_2$.

For example, assume $BCRT=5$ and $WCRT=100$, with a required delay of
$(200..200)$. Algorithm~\ref{alg:1}, results in $D$ being an empty set,
since $S_1=[2]$ and $S_2=[40]$, respectively. In such a case, the
resultant relaxation is: $40 \times 100$ using Algorithm~\ref{alg:2},
which results in a time of $4000$ units. Such large over-approximations
can be avoided by breaking the critical paths (worst case ticks) in the
program (by inserting pause constructs), just like in hardware design,
which is very \textit{unsurprising} considering Esterel has its origins
in hardware.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.45]{./tool.pdf}
  \label{fig:4}
  \caption{The tool-chain flow of SystemJ compiler with delays}
\end{figure}

@@

Our compiler has a switch that enables over-approximation as described
in Section~\ref{sec:extend-tehcn-vari} if the delay requirements are not
met. If this compiler switch is not enabled the compiler simply outputs
an error indicating the relaxation required. The system designer can
then either set the compiler switch of change the program by shortening
the critical path by introducing \texttt{pause} constructs. We show how
introducing \texttt{pause} constructs can help meet timing requirements
explicitly, using examples, later on in
Section~\ref{sec:experimental-results}.

>>>>

Figure~\ref{fig:5} pictorially represents the semantics of these delay
constructs. The construct \texttt{delay (M..N)} requires that upon
execution the control-flow is suspended for \textit{at least} $M$ units
and \textit{at most} $N$ units, depicted as bounded box in
Figure~\ref{fig:5}. There is no upper bound for the second delay variant
\texttt{delay(M)}, which requires that the control-flow suspends for
\textit{at least} $M$ units, but the following statement can be executed
after some countable delay, depicted with the unbounded box in
Figure~\ref{fig:5}. Finally, \texttt{delay (M..M)} requires that there
is an \textit{exact} delay of $M$ units, i.e., the control-flow suspends
execution for \textit{at least} and \textit{at most} $M$ units.

@@

This algorithm results in the \underline{smallest} relaxation required
for the real-time delay to be satisfied. The algorithm takes as input
the WCRT and set $S_2$, recall that set $S_2$ represents the logical
ticks required to satisfy the real-time requirement at the BCRT. We take
the very first value from set $S_2$ and multiply it with the WCRT to get
the relaxation $N$. The first element of set $S_2$ is returned as the
logical tick delay \texttt{d}. The fundamental observation is that we
have to delay for at least $M$ units of real-time, hence,
under-approximation is out of question. We can still over-approximate,
but to reduce the resultant error, we should over-approximate by least
possible value, which is the lower bound of set $S_2$. Thus, the lower
bound is considered to be the only element shared between the two sets
$S_1$ and $S_2$ and accordingly, the upper real-time bound is relaxed by
the multiplication of WCRT and the first element of set $S_2$.

For example, assume $BCRT=5$ and $WCRT=100$, with a required delay of
$(200..200)$. Algorithm~\ref{alg:1}, results in $D$ being an empty set,
since $S_1=[2]$ and $S_2=[40]$, respectively. In such a case, the
resultant relaxation is: $40 \times 100$ using Algorithm~\ref{alg:2},
which results in a time of $4000$ units. Such large over-approximations
can be avoided by breaking the critical paths (worst case ticks) in the
program (by inserting \texttt{pause} constructs), just like in hardware
design, which is very \textit{unsurprising} considering Esterel has its
origins in hardware.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.45]{./tool.pdf}
  \caption{The tool-chain flow of SystemJ compiler with delays}
  \label{fig:4}
\end{figure}

@@

Our compiler has a switch that enables over-approximation as described
in Section~\ref{sec:extend-tehcn-vari} provided the delay requirements
are not met. If this compiler switch is not enabled, the compiler simply
outputs an error indicating the relaxation required. The system designer
can then either set the compiler switch or change the program by
shortening the critical path by introducing \texttt{pause}
constructs. We show how introducing \texttt{pause} constructs can help
reduce the WCRT in Section~\ref{sec:experimental-results}.

### 2013-11-29 discussion.tex
Updated the paper

We dedicate this section to discuss the details of the delay semantics,
the dependence on time analyzable platforms, and the various uses and
programming paradigms that can be built using the $delay(M..N)$
construct.

@@

Many real-world systems require non-deterministic timing
constructs. Where the exact real-time delay is not known or should not
be known apriori. One such system is our motivating example -- the human
response time system (HRCTS) another one we presented was a real
printer-spooler embedded controller in
Section~\ref{sec:experimental-results}.

@@

In the above case, the program waits for signal \texttt{A} from the
environment for \textit{at\_least} \texttt{1 ms}, but it may wait longer
if other reactions/CDs are scheduled for execution. This second form of
delay statement us very useful for low priority tasks.

@@

The above program will emit signal \texttt{S} every \texttt{1 ms}. A
periodic reaction (or CD) requires special consideration, like any other
real-time periodic task. One should be aware of the interaction delays
and other \texttt{pause} constructs in the body of the reaction (or
CD). For example, consider the (perceived) periodic program and its
rewrite below.

@@

The system designer needs to be aware that just introducing a delay
construct does not make the reaction (or CD) periodic. The program on
the left does \textit{not} emit signal \texttt{S} every \texttt{1 ms},
because of the extra \texttt{pause} statement. Introducing additional
pauses introduces more logical ticks consequently increasing the
real-time delay.

@@

As mentioned previously, the proposed solution might need
over-approximation techniques (Algorithm~\ref{alg:2}) to guarantee
timing requirements (cf. Section~\ref{sec:extend-tehcn-vari}). This
might give the readers an impression that timer based solution proposed
for standard real-time operating systems can be utilized for
GALS/synchronous programs as well. This is a fallacy. Bourke et
al~\cite{Bourke2009a} have investigated exactly this technique
unsuccessfully. We elaborate upon the interplay of timers and reactivity
to make the problem of using timers explicit.

Preemption plays an important role in reactive languages. One needs to
carefully consider the interplay of \texttt{delay} semantics with the
preemption semantics of reactive languages. Previous attempts at
incorporating delays (using external timers) have only been partially
successful, because of the complex interplay between real-time and
preemption. Consider the simple example below, which models real-time
using external timers as in~\cite{rsh94}. Furthermore, synchronous/GALS
languages have always been targeted at real-time analyzable platforms
and constrained
environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}, one can use
our technique within such a setting with ease.

@@

As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
with the external timer. The above program sends a signal to an external
timer and waits for 10 ms to pass by. Consider what happens when signals
\texttt{S} and \texttt{TIMER} occur in the same logical tick, the
\texttt{await} statement is never executed (due to suspend) and hence,
we enter a deadlock. Such problems are completely avoided in our
technique, because we convert the real-time delays into logical delays
(\texttt{pause} constructs), which interact well with preemption.

@@

\subsubsection{Interaction of synchronous parallelism and delay}
\label{sec:inter-synchr-parall}

{\color{red}Maybe we should talk about how the control flow in other
  reactions and CDs is not affected by the delay statement.}

>>>>

We dedicate this section to discuss the details of the delay semantics,
the dependence on time analyzable platforms, and the various uses and
programming paradigms that can be built using the \texttt{delay (M..N)}
construct.

@@

Many real-world systems require non-deterministic timing
constructs. Where the exact real-time delay is not known or should not
be known apriori. One such system is our motivating example -- the human
response time system (HRCTS). Another one we presented was a real
printer-spooler embedded controller in
Section~\ref{sec:experimental-results}.

@@

In the above case, the program waits for signal \texttt{A} from the
environment for \textit{at\_least} \texttt{1 ms}, but it may wait longer
if other reactions/CDs are scheduled for execution. This second form of
delay statement is very useful for programming low priority tasks.

@@

The above program will emit signal \texttt{S} every \texttt{1 ms}. A
periodic reaction (or CD) requires special consideration, like any other
real-time periodic task. One should be aware of the interaction of
delays and other \texttt{pause} constructs in the body of the reaction
(or CD). For example, consider the (perceived) periodic program and its
rewrite below.

@@

The system designer needs to be aware that just introducing a delay
construct does not make the reaction (or CD) periodic. The program on
the left does \textit{not} emit signal \texttt{S} every \texttt{1 ms},
because of the extra \texttt{pause} statement. Introducing additional
pauses introduces more logical ticks, consequently increasing the
real-time delay.

@@

As mentioned previously, the proposed solution might need
over-approximation techniques (Algorithm~\ref{alg:2}) to guarantee
timing requirements (cf. Section~\ref{sec:extend-tehcn-vari}). This
might give the readers an impression that timer based solution proposed
for standard real-time operating systems should be utilized for
GALS/synchronous programs. This is a fallacy. Bourke et
al~\cite{Bourke2009a} have investigated this technique
unsuccessfully. We elaborate upon the interplay of timers and reactivity
to make the problem of using timers explicit.

Preemption plays an important role in reactive languages. One needs to
carefully consider the interplay of \texttt{delay} semantics with the
preemption semantics of reactive languages. Previous attempts at
incorporating delays (using external timers) have only been partially
successful, because of the complex interplay between real-time and
preemption. Consider the simple example below, which models real-time
using external timers as in~\cite{rsh94}.

@@

As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
with the external timer. The above program sends a signal to an external
timer and waits for 10 ms to pass by. Consider what happens when signals
\texttt{S} and \texttt{TIMER} occur in the same logical tick, the
\texttt{await} statement is never executed (due to suspend) and hence,
we enter a deadlock. Such problems are completely avoided in our
technique, because we convert the real-time delays into logical delays
(\texttt{pause} constructs), which interact well with preemption.
Furthermore, synchronous/GALS languages have always been targeted at
real-time analyzable platforms and constrained
environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}, one can use
our technique within such a setting with ease.

@@

% \subsubsection{Interaction of synchronous parallelism and delay}
% \label{sec:inter-synchr-parall}

% {\color{red}Maybe we should talk about how the control flow in other
%   reactions and CDs is not affected by the delay statement.}

### 2013-11-29 exps.tex
Updated text and added discussion text
Updated the paper

{\renewcommand{\arraystretch}{0.9}
\begin{table}[t]
	\small
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{delay(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{delay(2.4..2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 & \texttt{delay(1.667..1.667 ms)}      &                    &                      \\ 
											 & \texttt{delay(0.05..0.05 ms)}       &                    &                      \\ 
											 & \texttt{delay(0.3..0.3 ms)}        &                    &                      \\ 
											 & \texttt{delay(0.733..0.733 ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{delay(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{delay(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{delay(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{delay(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{delay(10000..10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\vspace{-20pt}
\end{table}
}

{\renewcommand{\arraystretch}{0.6}
\begin{table}[t]
\twocolumn[
\centering
\begin{@twocolumnfalse}
	\caption{Actual delays obtained for \texttt{delay} constructs in the
benchmark programs based on their BCRT and WCRT}
\label{fig:comparison}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0334 ms	& 0.0361 ms & 0.009613 ms &
	0.0263 ms & 0.1683 ms & 0.1881 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.1120 ms & 0.1116 ms & 0.0603 ms
	& 0.0491 ms & 0.8148 ms & 1.0115 ms\\ 
	\cmidrule(r){1-8}

	\multirow{2}{*}{Delay 1} & M..N & 50.3-200.3 ms	& 2.4-8.4882 ms	  &
	2.7-16.9378 ms & 1.2-30.3 ms & 10000-42483.2198 ms& 10000-53882.0045 ms \\
	\cmidrule(r){2-8}
	& d &1506-1888 & 68 & 281 & 46-617 & 59428 &53160\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 2} & M..N & N/A & 1.668-5.2452 ms & N/A &
	5.6-100.2 ms & N/A &10000-53882.0045 ms\\ 
	\cmidrule(r){2-8}
	& d & N/A & 48 & N/A & 213-2041 & N/A &53160\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 3} & M..N & N/A & 0.05-0.2232 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 2 & N/A & N/A & N/A &N/A\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 4} & M..N & N/A & 0.3-1.0044 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 9 & N/A & N/A & N/A &N/A\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 5} & M..N & N/A & 0.833-2.3436 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 21 & N/A
	& N/A & N/A &N/A\\ \bottomrule
\end{tabular}
\end{@twocolumnfalse}
]
\end{table}
}

@@

HRTCS is the human response system described in
Section~\ref{sec:motivating-example}. AECS (Access and Environmental
Control System) is an application that controls an intelligent room
environment described in~\cite{aecs_ispa}.  The system interacts with
two external timers through input/output signals in order to measure how
long the entrance door to the room is being opened and time the illegal
presence (e.g. intruder) that is being detected. The program has been
modified such that the delay constructs are used instead of external
timers as shown in Figure~\ref{delaycode}. In the original AECS, the
system needed to trigger external timer by emitting the output signal in
order to wait for a specific amount of time (e.g.
\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
expires, it notifies back to the system through the input signal
\texttt{TimeOut} which is polled at the \texttt{await} statement. With
the new delay constructs, those timers are not necessary and delays can
be expressed directly in the code as shown in Figure~\ref{delaycode:b}.

\begin{figure}[h!]
	\begin{SubFloat}{\label{delaycode:a}}
		\centering
		\begin{minipage}[b]{0.3\linewidth}
			\scriptsize
\begin{verbatim}
emit TimerTriggered;
trap(T){
  abort(DoorOpened){
    await(TimeOut);
    exit(T); 
  }
}
\end{verbatim}
		\end{minipage}
	\end{SubFloat}
\hspace{1cm}%
	\begin{SubFloat}{\label{delaycode:b}}
		\centering
		\begin{minipage}[b]{0.5\linewidth}
			\scriptsize
\begin{verbatim}
// No emit statement
trap(T){
  abort(DoorOpened){
    delay(10000 ms);
    exit(T);  
  }
}
\end{verbatim}
		\end{minipage}
	\end{SubFloat}
\caption{Replacing external timer with delay construct}
\label{delaycode}
\end{figure}

Motor is a stepper motor controller from~\cite{Bourke2009a}, used for
producing monochrome images on paper by applying a correct application
of current to a print head (a row of resistors). In this example,
duration of current applied is controlled by the feedback logic using
real-time delays which are expressed in Esterel constructs. We have
rewritten the program in SystemJ and compared between desired and
relaxed delays according to WCRT and BCRT of the program. Printer is an
example from~\cite{Schneider:1999:CRT:555233}, which models a printer and
a spooler in timed CSP.  Their operation is rather simple: as soon as
the spooler receives a print job, passes it to the printer through a
channel where actual printing process take place. The example
demonstrates that by specifying delay bounds (e.g. 1 to 5 time units) on
receiving and sending channel ends data exchange will occur as soon as
both processes become ready.

@@

HRTCS consists of 2 CDs, but only one with real-time delays. AECS also
has 2 CDs; CD1 consist of a single delay statement, while CD2 has 2
delay statements. In this case, as explained previously, delay
statements replace the external timers and corresponding interface (I/O)
signals. The system triggers an alarm 10 seconds after detection of an
intruder or generates a beep sound when an entrance door is being opened
for the same amount of time. In Printer, each spooler and printer
process is mapped to a single clock-domain. Spooler CD sends a job to
the Printer CD between 2.7 to 10.6 ms after receiving it from the
enviornment (e.g. a user). On the other hand, Printer CD notifies
environment that it starts printing between 1.2 to 30.3 ms once it
receives the job from the Spooler CD. Actual print job is started
between 5.6 to 100.2 ms after the notification. In this example, bounded
delay ensures that both channel communication and print job should occur
within the specified amount of time. Lastly, Motor is a synchronous
program with 5 delay statements where they controls amount of current
into the stepper motor coils.

@@

Table~\ref{fig:comparison} shows the result of our algorithm applied to
each benchmark program. Note that we allowed the compiler to relax the
upper real-time bounds even for bounded (\texttt{M..N}) or exact
(\texttt{M..M}) delays when it fails to find a feasible solution.
However, a user can prohibit this behavior through configuration that
such case is an error.

As a first step, BCRT and WCRT were calculated for all examples. Given
the real-time \mbox{\texttt{delay (M..N)}} provided by programmers,
required logical delay \textit{d} was determined by applying
Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
50.3..200.3 (ms) is only satisfied when \emph{d} is between 1506 and
1787 (as shown previously in
Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}). Similarly,
Printer/CD2 satisfies real-time ranges of 1.2..30.3 (ms) and 5.6..100.2
(ms) when \emph{d} is 46..617 and 213..2041 respectively. It is then
compiler's job to statically choose one of the values in this range.

% The upper bound \texttt{N} needed to be relaxed for all examples other
% than HRTCS to obtain a feasible \texttt{d}. Therefore,
On the other hand, Algorithm \ref{alg:2} is applied to those examples,
where relaxation was required, to determine the minimum upper real-time
bound \emph{N} such that there exist a solution for \emph{D} in
Algorithm \ref{alg:1}. This algorithm gives a unique number for logical
tick delay \emph{d} such that \emph{d} = Min(\emph{$S_2$}) =
Max(\emph{$S_1$}). The compiler, therefore, will choose \emph{d} based
on the newly computed \emph{N} for the examples: Printer/CD1, Motor and
AECS.

@@

The previous results show that HRTCS and Printer/CD2 were the only
examples that the compiler was able to find \textit{d} without relaxing
upper bound of \texttt{delay}. Figure~\ref{fig:graph} shows the accuracy
of our result by comparing differences between programmer specified
upper bounds and the actual bounds obtained from our algorithm.

@@

There are two main reasons that influence size of the relaxation: 1)
execution speed of a target platform 2) and the nature of the program
which runs on that specific platform. As one can easily speculate, clock
frequency of the processor directly impacts on the value of WCRT. On the
other hand, programmer can write a program such that large data
computation in a single tick increases WCRT of the program. Assuming we
do not change the architecture of the execution platform (i.e. JOP), the
difference factor can only be reduced by modifying the benchmark
programs by dividing the single longest tick, that corresponds to WCRT,
into a number of smaller ticks.

@@

% As one can see in Figure \ref{fig:comparison}, BCRT and WCRT of overall programs
% are smaller when they run on TP-JOP. For example, WCRT and BCRT of HRTCS are
% \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for TP-JOP (0.0028,
% 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the other hand, required
% logical delays \emph{d} are generally bigger on TP-JOP e.g. \(\times\)4.5 -
% \(\times\)5.9 and \(\times\)30.9 in Motor and Robot respectively. It is expected
% as their BCRT and WCRT differences are bigger on TP-JOP.
% This also led to greater increase in minimum upper real-time bounds \emph{N} for
% TP-JOP in every example.
% In AECS, there are two clock-domains using delay constructs that each has their
% own BCRT and WCRT. Again, individual BCRT and WCRT is smaller whereas \emph{d}
% is bigger on TP-JOP. One important property to note here, is that BCRT and WCRT
% of any clock-domains are invariant of \emph{d} as explained in Section
% \ref{sec:intr-real-time}. It is our fundamental assumption to find \emph{d} in
% our benchmark programs.
% >>>>>>> 4bffc4b1a43388f5ebf6cc737b9d3b246ee77dc7

>>>>

@@

HRTCS is the human response system described in
Section~\ref{sec:motivating-example}. AECS (Access and Environmental
Control System) is an application that controls an intelligent room
environment described in~\cite{aecs_ispa}. The system interacts with two
external timers through input/output signals in order to measure how
long the entrance door to the room is being opened to detect intruders.
The program has been modified such that the delay constructs are used
instead of external timers as shown in Figure~\ref{delaycode}. In the
original AECS, the system needed to trigger external timer by emitting
the output signal in order to wait for a specific amount of time (e.g.
\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
expires, it notifies the system through the input signal
\texttt{TimeOut}, which is polled using the \texttt{await}
statement. With the new delay constructs, these timers are not necessary
and delays can be expressed directly in the code as shown in
Figure~\ref{delaycode:b}.

\begin{figure}[h!]
  \begin{SubFloat}{\label{delaycode:a}Original program}
    \centering
    \begin{minipage}[b]{0.3\linewidth}
      \scriptsize
\begin{verbatim}
emit TimerTriggered;
trap(T){
  abort(DoorOpened){
    await(TimeOut);
    exit(T); 
  }
}
\end{verbatim}
    \end{minipage}
  \end{SubFloat}
  \hspace{1cm}%
  \begin{SubFloat}{\label{delaycode:b}Program with delay constructs}
    \centering
    \begin{minipage}[b]{0.5\linewidth}
      \scriptsize
\begin{verbatim}
// No emit statement
trap(T){
  abort(DoorOpened){
    delay(10000 ms);
    exit(T);  
  }
}
\end{verbatim}
		\end{minipage}
	\end{SubFloat}
\caption{Replacing external timer with delay construct}
\label{delaycode}
\end{figure}

Motor is a stepper motor controller from~\cite{Bourke2009a}, used for
producing monochrome images on paper by applying a correct application
of current to a print head (a row of resistors). In this example,
duration of the current applied is controlled by the feedback logic
using real-time delays, which are expressed using reactive
constructs. Printer is an example from~\cite{Schneider:1999:CRT:555233},
which models a printer and a spooler in timed CSP.  Their operation is
rather simple: as soon as the spooler receives a print job, passes it to
the printer through a channel where actual printing process take
place. This example along with HRCTS demonstrates the use of
non-deterministic delays.

@@

HRTCS consists of 2 CDs, but only one with real-time delays. AECS also
has 2 CDs; CD1 consist of a single delay statement, while CD2 has 2
delay statements. In this case, as explained previously, delay
statements replace the external timers and corresponding interface (I/O)
signals. The system triggers an alarm 10 seconds after detection of an
intruder or generates a beep sound when an entrance door is being opened
for the same amount of time. In the Printer example, spooler and printer
processes are mapped to separate CDs. Spooler CD sends a job to the
Printer CD between 2.7 to 10.6 ms after receiving it from the
environment (e.g. a user). On the other hand, Printer CD notifies
environment that it starts printing between 1.2 to 30.3 ms once it
receives the job from the Spooler CD. Actual print job is started
between 5.6 to 100.2 ms after the notification. In this example, bounded
delay ensures that both channel communication and print job should occur
within the specified amount of time. Lastly, Motor is a synchronous
program with 5 delay statements that control the amount of current
applied to the stepper motor coils.

@@

Table~\ref{fig:comparison} shows the result of our algorithm applied to
each benchmark program. Note that we allowed the compiler to relax the
upper real-time bounds even for bounded (\texttt{M..N}) or exact
(\texttt{M..M}) delays when it failed to find a feasible solution.
However, a user can prohibit such behavior using compiler switch as
described in Section~\ref{sec:tool-chain-flow}.

As a first step, BCRT and WCRT were calculated for all examples. Given
the real-time \mbox{\texttt{delay (M..N)}} provided by programmers,
required logical delay \textit{d} was determined by applying
Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
50.3..200.3 (ms) is only satisfied when \emph{d} is between 1506 and
1787 (as shown previously in
Section~\ref{sec:find-logic-delay}). Similarly, Printer/CD2 satisfies
real-time ranges of 1.2..30.3 (ms) and 5.6..100.2 (ms) when \emph{d} is
46..617 and 213..2041, respectively. It is then the compiler's job to
statically choose one of the values in this range.

% The upper bound \texttt{N} needed to be relaxed for all examples other
% than HRTCS to obtain a feasible \texttt{d}. Therefore,
On the other hand, Algorithm \ref{alg:2} is applied to those examples,
where relaxation was required, to determine the minimum upper real-time
bound \emph{N} such that there exists a solution for \emph{D} in
Algorithm \ref{alg:1}. This algorithm gives a unique number for logical
tick delay \emph{d} such that \emph{d} = Min(\emph{$S_2$}) =
Max(\emph{$S_1$}). The compiler, therefore, will choose \emph{d} based
on the newly computed \emph{N} for the examples: Printer/CD1, Motor and
AECS.

@@

The previous results show that HRTCS and Printer/CD2 were the only
examples that the compiler was able to find \textit{d} without relaxing
upper bound of \texttt{delay}. Figure~\ref{fig:graph} shows the accuracy
of our results by comparing differences between programmer specified
upper bounds (Table~\ref{fig:benchlist}) and the actual bounds obtained
using our techniques (Table~\ref{fig:comparison}).

@@

There are two influential factors in determining the relaxation: 1)
execution speed of a target platform and 2) the nature of the program
that runs on this platform. Clock frequency of the processor directly
impacts the value of WCRT. On the other hand, programmer can write a
program such that large data computations in a single logical tick
increase WCRT of the program. Assuming we do not change the architecture
of the execution platform (i.e. JOP), the difference factor can only be
reduced by modifying the benchmark programs such that the critical paths
are broken up using \texttt{pause} constructs, thereby reducing the
WCRT.

@@

% As one can see in Figure \ref{fig:comparison}, BCRT and WCRT of overall programs
% are smaller when they run on TP-JOP. For example, WCRT and BCRT of HRTCS are
% \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for TP-JOP (0.0028,
% 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the other hand, required
% logical delays \emph{d} are generally bigger on TP-JOP e.g. \(\times\)4.5 -
% \(\times\)5.9 and \(\times\)30.9 in Motor and Robot respectively. It is expected
% as their BCRT and WCRT differences are bigger on TP-JOP.
% This also led to greater increase in minimum upper real-time bounds \emph{N} for
% TP-JOP in every example.
% In AECS, there are two clock-domains using delay constructs that each has their
% own BCRT and WCRT. Again, individual BCRT and WCRT is smaller whereas \emph{d}
% is bigger on TP-JOP. One important property to note here, is that BCRT and WCRT
% of any clock-domains are invariant of \emph{d} as explained in Section
% \ref{sec:intr-real-time}. It is our fundamental assumption to find \emph{d} in
% our benchmark programs.
% >>>>>>> 4bffc4b1a43388f5ebf6cc737b9d3b246ee77dc7
{\renewcommand{\arraystretch}{0.9}
\begin{table}[t]
	\small
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{delay(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{delay(2.4..2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 & \texttt{delay(1.667..1.667 ms)}      &                    &                      \\ 
											 & \texttt{delay(0.05..0.05 ms)}       &                    &                      \\ 
											 & \texttt{delay(0.3..0.3 ms)}        &                    &                      \\ 
											 & \texttt{delay(0.733..0.733 ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{delay(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{delay(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{delay(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{delay(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{delay(10000..10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\vspace{-20pt}
\end{table}
}

{\renewcommand{\arraystretch}{0.6}
\begin{table}[t!]
\twocolumn[
\centering
\begin{@twocolumnfalse}
	\caption{Actual delays obtained for \texttt{delay} constructs in the
benchmark programs based on their BCRT and WCRT}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0334 ms	& 0.0361 ms & 0.009613 ms &
	0.0263 ms & 0.1683 ms & 0.1881 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.1120 ms & 0.1116 ms & 0.0603 ms
	& 0.0491 ms & 0.8148 ms & 1.0115 ms\\ 
	\cmidrule(r){1-8}

	\multirow{2}{*}{Delay 1} & M..N & 50.3-200.3 ms	& 2.4-8.4882 ms	  &
	2.7-16.9378 ms & 1.2-30.3 ms & 10000-42483.2198 ms& 10000-53882.0045 ms \\
	\cmidrule(r){2-8}
	& d &1506-1888 & 68 & 281 & 46-617 & 59428 &53160\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 2} & M..N & N/A & 1.668-5.2452 ms & N/A &
	5.6-100.2 ms & N/A &10000-53882.0045 ms\\ 
	\cmidrule(r){2-8}
	& d & N/A & 48 & N/A & 213-2041 & N/A &53160\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 3} & M..N & N/A & 0.05-0.2232 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 2 & N/A & N/A & N/A &N/A\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 4} & M..N & N/A & 0.3-1.0044 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 9 & N/A & N/A & N/A &N/A\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 5} & M..N & N/A & 0.833-2.3436 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 21 & N/A
	& N/A & N/A &N/A\\ \bottomrule
\end{tabular}
\label{fig:comparison}
\end{@twocolumnfalse}
]
\end{table}
}
